---
title: "Untitled"
author: "Akshat Jain,Roll no. 204161003"
date: "4/17/2021"
output:
  pdf_document: default
  powerpoint_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Motivation:
To study the production of major crops in various states of India in the last 50 years(from 1966 to 2017) and visualize some pattern between them.

Agriculture is the most important sector of Indian Economy. Indian agriculture sector accounts for 18 per cent of India's gross domestic product (GDP) and provides employment to 50% of the countries workforce.Despite this,it is considered highly inefficient, wasteful, and incapable of solving the hunger and malnutrition problems.
The use of modern technology such as automated machines, crop sensors, etc. can significantly improve the agriculture sector.

Dataset Source: http://data.icrisat.org/dld/src/crops.html ICRISAT-District Level Data.csv
Sector: Agriculture
Sub-sector: Crops production
Target states/UTs: 20 major states
#Objective
1. states which contribute to maximum crop production in quantity for year 2017
2. states which have maximum agricultural land for producing crops
3. comparing the production of crops from 1966 to 2017
4. comparing the area under production from 1966 to 2017
5. relation between quantity of production and area of production
6. crops which are produced in maximum quantity
7. crops which are grown in maximum area and analysing their trend,discussing on ways to improve the agriculture sector



```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(reshape2)
library(scales)
library(plotly)
library(treemapify)
data <- read.csv("ICRISAT-District Level Data.csv", header = TRUE)
data[data < 0] <- NA
data<-na.locf(na.locf(data), fromLast = TRUE)
data$total_production <- rowSums(data[,c(7,10,13,16,19,22,25,28,31,34,37,40,43,46,49,52,55,58,61,64,67,70,73)])
```
#1
1. states which contribute to maximum crop production in quantity for the year 2017

Plot used: Bar plot, TreeMap, Pie chart, Proportional Symbol map
Considering total crop production quantity:
1. Bar plot
```{r}

newdata <- data[ which(data$Year=='2017'),]
ggplot(data = newdata, aes(x = reorder(State.Name,-newdata$total_production),y= newdata$total_production))+
  geom_bar(aes(fill= State.Name),stat = "identity",size = 2) +
  ggtitle("Comparision of total crop production between 
              Indian states for the year 2017")+
   labs(y=" Total production in 1000 metric tons", x = "States Name")+
  theme_bw() +
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```
2. TreeMap
```{r}
data$total_area <- rowSums(data[,c(6,9,12,15,18,21,24,27,30,33,36,39,42,45,48,51,54,57,60,63,66,69,72,75,76,77,78,79,80)], na.rm=TRUE)

newdata <- data[ which(data$Year=='2017'),]

dat <-newdata %>% 
  group_by(State.Name) %>% 
   summarise(across(c(total_production, total_area), sum) )
ggplot(dat, aes(area = total_area, fill = total_production, label = State.Name)) +
geom_treemap() +
geom_treemap_text(fontface = "italic", colour = "black", place = "centre", grow = TRUE) + 
  scale_fill_gradient(low="white", high="darkgreen")
```

3. Pie Chart
```{r}
newdata <- data[ which(data$Year=='2017'),]
fig <- plot_ly(newdata, labels = ~State.Name,
               values = newdata$total_production, type = 'pie',
               textposition = 'inside',
               textinfo = 'label+percent',
               insidetextfont = list(color = '#FFFFFF'),
               marker = list(line = list(color = 'black', width = 1)),
               showlegend = FALSE)


fig %>%
  layout(title = 'Percentage share of state in total Production of crop for the year 2017')
```
Observations:
Uttar Pradesh contributed the maximum crop production for the year 2017 followed by Madhya Pradesh and Punjab.
4. Proportional symbol map:
```{r}
library(rgdal)
library(sp)
shp = readOGR(dsn = ".",layer = "IND_adm0")
plot(shp)
states <- fortify(shp)

```
I tried merging the shape file with csv data file, but could not do it because of having only 20 states in the csv data file.


Considering production of rice for the year 2017:
1. Bar Plot:
```{r}
ggplot(data = newdata, aes(x = reorder(State.Name,-RICE.PRODUCTION..1000.tons.),y = RICE.PRODUCTION..1000.tons., fill = State.Name))+
  geom_bar(stat = "identity",size = 2) +
  ggtitle("Comparision of total rice production between 
              Indian states for the year 2017")+
   labs(y=" Total production in 1000 metric tons", x = "States Name") +
  theme_bw() +
  theme(legend.position = "none")+
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
2. Pie Chart
```{r}
newdata <- data[ which(data$Year=='2017'),]
fig <- plot_ly(newdata, labels = ~State.Name,
               values = newdata$RICE.PRODUCTION..1000.tons, type = 'pie',
               textposition = 'inside',
               textinfo = 'label+percent',
               insidetextfont = list(color = '#FFFFFF'),
               marker = list(line = list(color = 'black', width = 1)),
               showlegend = FALSE)


fig %>%
  layout(title = 'Percentage share of state in Production of rice
                            for the year 2017')
```
Observation:
West Bengal produced maximum rice quantity followed by Uttar Pradesh and Punjab

Pie chart is more convenient for comparing the quantity of crop production among various states


Considering production of wheat for the year 2017:
Bar Plot:
```{r}
ggplot(data = newdata, aes(x = reorder(State.Name,-WHEAT.PRODUCTION..1000.tons.),y = WHEAT.PRODUCTION..1000.tons., fill = State.Name))+
  geom_bar(stat = "identity",size = 2) +
  ggtitle("Comparision of total wheat production between 
              Indian states for the year 2017")+
   labs(y=" Total production in 1000 metric tons", x = "States Name") +
  theme_bw() +
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Pie Chart

```{r}
newdata <- data[ which(data$Year=='2017'),]
fig <- plot_ly(newdata, labels = ~State.Name,
               values = newdata$WHEAT.PRODUCTION..1000.tons., type = 'pie',
               textposition = 'inside',
               textinfo = 'label+percent',
               insidetextfont = list(color = '#FFFFFF'),
               marker = list(line = list(color = 'black', width = 1)),
               showlegend = FALSE)


fig %>%
  layout(title = 'Percentage share of states in wheat Production 
         of crop for the year 2017')
```
Observation:
Uttar Pradesh produced the highest wheat quantity for the year 2017, contributing nearly 1/3rd of the total wheat production. It is followed by Madhya Pradesh and Punjab.


#2.
states which have maximum agricultural land for producing crops

Bar plot:
```{r}
newdata <- data[ which(data$Year=='2017'),]
ggplot(data = newdata, aes(x = reorder(State.Name,-newdata$total_area),y= newdata$total_area))+
  geom_bar(aes(fill= State.Name),stat = "identity",size = 2) +
  ggtitle("Comparision of total area under crop production between 
              Indian states for the year 2017")+
   labs(y=" Total area in  Hectares", x = "States Name")+
  theme_bw() +
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```
Pie chart
```{r}
newdata <- data[ which(data$Year=='2017'),]
fig <- plot_ly(newdata, labels = ~State.Name,
               values = newdata$total_area, type = 'pie',
               textposition = 'inside',
               textinfo = 'label+percent',
               insidetextfont = list(color = '#FFFFFF'),
               marker = list(line = list(color = 'black', width = 1)),
               showlegend = FALSE)


fig %>%
  layout(title = '% share of state in total area under Production of crop for the year 2017')
```
Observation:
Rajasthan has maximum area under production of crops followed by Madhya Pradesh and Maharashtra
#3.
Comparing the total crop production of crops yearwise from 1966 to 2017
Bar Plot:
```{r}
data$total_production <- rowSums(data[,c(7,10,13,16,19,22,25,28,31,34,37,40,43,46,49,52,55,58,61,64,67,70,73)], na.rm=TRUE)

ggplot(data = data, aes(x = Year, y = total_production, fill = Year))+
  geom_bar(stat = "identity",position="stack") + 
  scale_y_continuous(name="Total production of crops in tons", labels = comma)+
  ggtitle("Comparing the total crop production of crops yearwise
          from 1966 to 2017")+
  theme_bw()

```
Observation:
Total crop production has increased almost 4 times from 1966 to 2017.


#4.
Comparing the area under production of crops yearwise from 1966 to 2017
```{r}
data$total_area <- rowSums(data[,c(6,9,12,15,18,21,24,27,30,33,36,39,42,45,48,51,54,57,60,63,66,69,72,75,76,77,78,79,80)], na.rm=TRUE)

ggplot(data = data, aes(x = Year, y = total_area, fill = Year))+
  geom_bar(stat = "identity",position="stack") + 
  scale_y_continuous(name="Total area under production of crops in hectare", labels = comma)+
  ggtitle("Comparing the area under production of crops yearwise
          from 1966 to 2017")+
  theme_bw()
```
Observation:
The area under production of crops increased gradually from 1966 to 2010 but has been declining rapidly since 2010. Urbanization and industrialization can be one of the reasons for its decline.


#5. 
relation between quantity of production and area of production
Plot: Scattered plot
```{r}
newdata <- data[ which(data$Year=='2017'),]

dat <-newdata %>% 
  group_by(State.Name) %>% 
   summarise(across(c(total_production, total_area), sum) )

ggplot(dat, aes(y = total_production, x = total_area, color = State.Name)) + geom_point(size = 4) +
  labs(y = "Production in 1000 tons", x = "Total area in hectares", 
   title = "Comparing production of crops with total area
       for various states for year 2017") +
  theme(plot.title = element_text(hjust = 0.5, size = 13))
```
observation:
Production of crops increases almost linearly with the increase in area.States such as Rajasthan, Odisha have very low yield output.
Punjab has very high yield output because of use of modern techniques

#6. 
major crops produced and their trend
Plot: Stacked bar plot
```{r, fig.height = 10, fig.width = 15}
dfm <- melt(data[,c(2,7,10,40,67,70,73)],id.vars = 1)
ggplot(dfm,aes(x = Year,y = value)) + 
    geom_bar(aes(fill = variable),stat = "identity",position = "stack")+

     scale_y_continuous(name="Production in Tons", labels = comma)+
  theme(legend.position="bottom")+
  scale_fill_discrete(name = "Crop", labels = c("Rice",     "Wheat","pulses","oilseeds", "Sugarcane","Cotton"))+
  ggtitle("Comparing various crops production from 1966 to 2017")+
  theme(text = element_text(size = 25))  
```

Observation: It is seen that there is continuous rise in the production of staple crop like wheat and rice. On the other hand, production of oil seeds and pulses has been declining gradually. Production of sugarcane and cotton has been constant over the years.

#7.
Comparing area under various crop production from 1966 to 2017

```{r, fig.height = 10, fig.width = 15}
dfm <- melt(data[,c(2,6,9,39,66,69,73)],id.vars = 1)
ggplot(dfm,aes(x = Year,y = value)) + 
    geom_bar(aes(fill = variable),stat = "identity",position = "stack")+

     scale_y_continuous(name="Area in Hectares", labels = comma)+
  theme(legend.position="bottom")+
  scale_fill_discrete(name = "Crop", labels = c("Rice",     "Wheat","pulses","oilseeds", "Sugarcane","Cotton"))+
  ggtitle("Comparing area under various crop production from 1966 to 2017")+
  theme(text = element_text(size = 25))  
```

Observation:
Area under wheat and rice has remain constant. There is sharp decline in the area under oilseeds in the last decade.

Conclusion:
One pattern is seen among the farmers from the above visualization. Most of the farmers rely on traditional crops like wheat and rice. They tend to avoid growing oilseeds mainly because of its low productivity yield and high competition from the international market. India imports large quantity of edible oil such as soyabean oil and Palm oil from neighbouring countries. 
Government can provide subsidies to the farmer for growing such crops which will help curb imports on such items and also provide farmers new opportunities.
This will be a win-win situation for both Indian Economy and the farmers.
